Criticisms of the IHME COVID-19 model

### 2020-05-12 text/20/ihme-covid-model.Rmd
Added bioconductor-library

---
title: "Criticisms of the IHME COVID-19 model"
author: "Steve Simon"
source: "New"
date: "2020-04-23"
category: Blog post
tags: Epidemiology
output: html_document
---

@@

@@

The alternative SEIR models generally use information from the daily infection rate to estimate parameters and buil predictions, though they can incorporate information from the

@@

>>>>

---
title: "Criticisms of the IHME COVID-19 model"
author: "Steve Simon"
source: "New"
category: Blog post
tags: Epidemiology
output: html_document
---

@@

This blog post was created on 2020-04-23 and last modified on `r Sys.Date()`.

@@

The alternative SEIR models generally use information from the daily infection rate to estimate parameters and build predictions, though they can incorporate information from the daily death rates as well. I have not seen any SEIR models that do this, but there are many SEIR models out there and I've only looked at a few.

@@

### Coda

I am not an expert in infectious disease epidemiology. In fact, I'm not much of an expert in anything. I am more or less a generalist. I know a little bit about a lot of different things. I can't really compete with teams of researchers who have been modeling epidemics for many decades. If I can contribute something, though, it is in a critical appraisal of the various prediction models that are out there for COVID-19.

A critical appraisal is not a criticism. It is not a listing of all the faults that a model might have. Rather, it is a description of the features that a model has. It explains the rationale for these features and contrasts them with alternatives. It presents both the weaknesses AND THE STRENGTHS of a model.

Everyone, including myself, gets trapped into a hypercritical approach to critical appraisal. The easiest thing in the world is to build a laundry list of faults. This can lead to a sense of cynicism towards all research, as all research has flaws. A good critical appraisal compares not to a standard of perfection in modeling, but rather to other practical modeling approaches.

I hope I have not been too hypercritical in my review of the IHME model. Unfortunately, if there is anything I've learned over the years, it is a sense of humility about my own limitations. The faults I point out about others are the faults I suffer from myself. I can't get rid of my faults, but I can try to mitigate them.

One of the sites listed above linked to an excellent article that got me to think a bit on the philosophical side of research.

Gog JR. How you can help with COVID-19 modelling. Nat Rev Phys, 2020-04-08. https://doi.org/10.1038/s42254-020-0175-7. Available in [html format](https://www.nature.com/articles/s42254-020-0175-7) and [pdf format](https://www.nature.com/articles/s42254-020-0175-7.pdf).
